%===================================== CHAP 1 =================================

%Only for right page counting---------------
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\chaptermark}[1]{\markboth{\chaptername\ \thechapter.\ #1}{}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\ #1}}
\renewcommand{\headrulewidth}{0.1ex}
\renewcommand{\footrulewidth}{0.1ex}
\fancyfoot[LE,RO]{\thepage}
\fancyhead[LE]{\leftmark}
\fancyhead[RO]{\rightmark}
\fancypagestyle{plain}{\fancyhf{}\fancyfoot[LE,RO]{\thepage}\renewcommand{\headrulewidth}{0ex}}

\pagenumbering{arabic} 				
\setcounter{page}{1}
%--------------------------------------------

\chapter{Introduction}\label{introduction}

For the past decade, Fantasy Sports has experienced an enormous growth, both by terms of active players and by media attention. Fantasy Sports is a type of online game where participants assemble imaginary teams consisting of real players within a professional sport. Fantasy teams compete in both private- and public leagues, where the teams gain points based on the statistical performance of real players in actual games. According to Fantasy Sports Trade Association (2017), it is estimated that 59.3 million players in the U.S and Canada are competing in some kind of Fantasy Sports. Fantasy Premier League (FPL) the Fantasy Sports of England's top division football league, Premier League. It is the largest Fantasy Sports in Europe, and has more than 5.5 million active players. 
\newpar
Fantasy Sports has its origin from the latter part of 1950s when a limited partner of the Oakland Raiders, Wilfred Winkenbach, invented a Fantasy Sports for golf. Each player selected a team of professional golfers, and the player who suggested the team with the lowest total score would win the tournament. In 1962, Winkenbach suggested the first fantasy league for American Football, initiating the birth of what has become a billion-dollar business. Nowadays, Fantasy Sports are available for several sports, including American football, basketball, ice hockey, baseball, football and cricket.  
\newpar
Typically, a Fantasy Sports team consists of selecting a squad consisting of a given amount of players in each position on the field. Each real-life player is given a buying price based on their skill level and ability to perform well in the Fantasy Sports. The performance of each player in the actual games is measured in terms of points, where an action like scoring a goal will provide the player with a specified amount of points. In order to avoid players from selecting only top rated real-life players, each manager is given a restricted budget. 
\newpar
For the past 10 years, a new phenomenon called Daily Fantasy Sports (DFS) has arisen. DFS is an alternate of the original Fantasy Sports, allowing players to compete for short time-periods, typically for one day or a weekend. DFS is structured in a way where managers pay a fee for participating in a competition and the winners receive a predetermined amount of the entry fees. In the U.S, FanDuel and DraftKings are the main providers of DFS, with both companies having a worth of more than \$1 billion. The two companies nearly collected a combined entry fee of \$3 billion in 2015. For instance, DraftKings arranged a 1-day contest for the Major League Baseball in 2015, with combined winning prizes of \$500 000 where the first-place finisher received \$125 000. 
\newpar
Although Daily Fantasy Sports are most common for American Sports, FanDuel added the English Premier League to its Daily Fantasy Sports roster in 2016. In 2017 Unibet, a Nordic bookmaker, launched its first Daily Fantasy Sports, Fantasy Sports Beta, allowing their customers to compete for money within the English Premier League. To our knowledge, this is the first European Daily Fantasy Sports for the Premier League. 
\newpar
Fantasy Premier League managers usually select their teams based on their subjective opinion, common knowledge, tips suggested by football experts, blogs and discussion forums. However, more dedicated players may use statistical analysis and put more effort in composing the best performing team. Since Fantasy Sports is relative small in Europe compared to the U.S, there is a lack of academic work performed on Fantasy Premier League.
\newpar
The main objective of this project is to create an optimization model that is competitive to skilled FPL managers. The model should make decisions of which players the team should be composed of, and hence regarded as the Fantasy Team Composition Problem (FTCP). As far as we know, there are not similar models produced, and we therefore find it interesting to see whether it is possible for an optimization model to outperform football experts' intuition and knowledge. 
\newpar
As there exist several optimization models for ice hockey \citep{drafting_hockey_pools} and American football\citep{Fry}, the main objective of this project is to produce a competitive model for Fantasy Premier League as well. Compared to the models suggested for the mentioned Fantasy Sports, there is a difference when modelling a Fantasy Premier League team. While the classical American Fantasy leagues consist of selecting a team for an entire season and not allowing for transfers, the Fantasy Premier League allows for making transfers of players during the season. Thus, FPL managers has to consider making new decisions each gameweek in a season. Therefore, this project propose a model for making optimal decisions in a multi-week manner, not only at the start of a season or for just one gameweek as in the Daily Fantasy Sports. To our knowledge, this is the first model suggested for making fantasy decisions in a multi-week manner.
\newpar
Further, the approach suggested can be generalized to apply for other situations facing decision-making similar to those in FPL. For instance, the team composition model may be of inspiration to actual football teams when considering which players to sign during a transfer window. Even though some adjustments should be made, the characteristics of the problems are related in a great manner. As for FPL, a football team has got a restricted budget when purchasing new players. While FPL managers base their transfers on projected points gained by a player, football managers purchase players they expect to perform well in the future. By further comparison, FPL consists of a set of players to select, while the actual clubs can buy players from a similar set of players, namely the set of all players available in a transfer window.


\newpage

The project is structured as follows. Chapter 2 provides the problem description. Here, the rules and point system of Fantasy Premier League are presented. Chapter 3 provides a structural literature review of Fantasy Sports optimization models, scoreline prediction and measuring of player performance. Chapter 4 describes the mathematical formulation of the problem. It includes the problem structure, assumptions as well as a mathematical model for the Fantasy Team Composition Problem. Chapter 5 describes the solution approach used in this project. Chapter 6 describes the experimental setup of the approach as well as it introduces the data sets used in order to suggest a solution. Chapter 7 provides the results obtained by the optimization model and compare them to the weekly rankings of FPL. Chapter 8 gives a conclusion of the work that has been done. Finally, Chapter 9 addresses the potential for further research.
